---
title: "Getting and Cleaning Data Final Assesment"
output: html_document
---
***
###Detailed instruction on how to use "run_analysis.R"
* Download and unzip the data from:
   https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip
   
* The data should be saved to a directory named "Dataset".

* The script "run_analysis.R" should be in the parent directory.

* Set the working directory of "RGui" or "RStudio" to parent directory of "run_analysis.R".

* Run the script with the command:  source("run_analysis.R") in R.

####Reading of raw data:
```{r}
#Collect raw data
trainData <- read.table(".\\Dataset\\train\\X_train.txt")
trainLabel <- read.table(".\\Dataset\\train\\y_train.txt")
trainSubject <- read.table(".\\Dataset\\train\\subject_train.txt")
testData <- read.table(".\\Dataset\\test\\X_test.txt")
testLabel <- read.table(".\\Dataset\\test\\y_test.txt") 
testSubject <- read.table(".\\Dataset\\test\\subject_test.txt")
```
####Dimensions of raw data:
```{r}
dim(trainData)
dim(trainLabel)
dim(trainSubject)
dim(testData)
dim(testLabel)
dim(testSubject)
```

####To merge the data sets, we use the rbind() function
```{r}
mergeData <- rbind(trainData, testData)
mergeLabel <- rbind(trainLabel, testLabel)
mergeSubject <- rbind(trainSubject, testSubject)
```

####Dimensions of the merged data sets:
```{r}
dim(mergeData)
dim(mergeLabel)
dim(mergeSubject)
```
####To subset the correct data, we must first obrain information from the "features.txt" file. This information is the used to created a subset of the data based on the correct features.
```{r}
features <- read.table(".\\Dataset\\features.txt")
index <- grep("mean\\(\\)|std\\(\\)", features[, 2])
mergeData <- mergeData[, index ]
```

####We then recreate headings for the data based on Tidy Data principles, removing obsolete, and capatilizing, characters.
```{r}
names(mergeData) <- gsub("\\(\\)", "", features[index, 2])
names(mergeData) <- gsub("mean", "Mean", names(mergeData))
names(mergeData) <- gsub("std", "Std", names(mergeData))
names(mergeData) <- gsub("-", "", names(mergeData)) 
activity <- read.table(".\\Dataset\\activity_labels.txt")
activity[, 2] <- tolower(gsub("_", "", activity[, 2]))
substr(activity[2, 2], 8, 8) <- toupper(substr(activity[2, 2], 8, 8))
substr(activity[3, 2], 8, 8) <- toupper(substr(activity[3, 2], 8, 8))
activityLabel <- activity[mergeLabel[, 1], 2]
mergeLabel[, 1] <- activityLabel
names(mergeLabel) <- "activity"
names(mergeSubject) <- "subject"

tidyData <- cbind(mergeSubject, mergeLabel, mergeData)
```
####The dimensions of the tidy data:
```{r}
dim(tidyData)
```
####The new tidy data set is now created and exported.
```{r}
write.table(tidyData , "tidy_data.txt")
```

####The tidy data set is now used as an input to calculate the mean of the tidy data.
```{r}
sLen <- length(table(mergeSubject))
aLen <- dim(activity)[1]
cLen <- dim(tidyData )[2]
means <- matrix(NA, nrow=sLen*aLen, ncol=cLen) 
means <- as.data.frame(means)
colnames(means) <- colnames(tidyData)
index <- 1
for(i in 1:sLen) {
    for(j in 1:aLen) {
        means[index, 1] <- sort(unique(mergeSubject)[, 1])[i]
        means[index, 2] <- activity[j, 2]
        t1 <- i == tidyData$subject
        t2 <- activity[j, 2] == tidyData$activity
        means[index, 3:cLen] <- colMeans(tidyData[t1&t2, 3:cLen])
        index <- index + 1
    }
}
```
####The dimensions of the means of tidy data:
```{r}
dim(means)
```
####The tidy data with means is now saved.
```{r}
write.table(means, "tidy_data_with_means.txt")
```


* After the script has completed, 2 additional files will have been written to the "data" folder, namely:
"tidy_data.txt"" and "tidy_data_with_means.txt".

The created files will contain all the data required for this assignment in a "Tidy Data" format.